---
title: "IPV Detection Experiment Quality Report"
subtitle: "Comprehensive Analysis of LLM Performance & Prompt Engineering"
author: "Automated Analysis System"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    theme: cosmo
    highlight: tango
    df_print: paged
    number_sections: true
---

<style>
.metric-box {
  background: #f8f9fa;
  border-left: 4px solid #007bff;
  padding: 15px;
  margin: 10px 0;
}
.warning-box {
  background: #fff3cd;
  border-left: 4px solid #ffc107;
  padding: 15px;
  margin: 10px 0;
}
.success-box {
  background: #d4edda;
  border-left: 4px solid #28a745;
  padding: 15px;
  margin: 10px 0;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 6,
  dpi = 300,
  fig.align = 'center'
)

# Load libraries
library(DBI)
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(ggplot2)
library(DT)
library(scales)

# Custom ggplot theme
theme_report <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0),
    plot.subtitle = element_text(size = 12, color = "gray30", hjust = 0),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold", size = 12)
  )
theme_set(theme_report)

# Helper functions
fmt_pct <- function(x) sprintf("%.1f%%", x * 100)
fmt_num <- function(x) format(round(x), big.mark = ",")
fmt_dec <- function(x, digits = 3) sprintf(paste0("%.", digits, "f"), x)

# Color palette
colors <- c(
  primary = "#007bff",
  success = "#28a745",
  warning = "#ffc107",
  danger = "#dc3545",
  info = "#17a2b8"
)
```

```{r load_data, include=FALSE}
# Read database credentials from .env
env_file <- "../.env"
if(file.exists(env_file)) {
  env_vars <- readLines(env_file)
  for(line in env_vars) {
    if(grepl("^PG_", line) && grepl("=", line)) {
      parts <- strsplit(line, "=", fixed = TRUE)[[1]]
      if(length(parts) >= 2) {
        env_name <- trimws(parts[1])
        env_value <- trimws(paste(parts[-1], collapse = "="))
        do.call(Sys.setenv, setNames(list(env_value), env_name))
      }
    }
  }
}

# Connect to PostgreSQL
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  host = Sys.getenv("PG_HOST", "memini.lan"),
  port = as.integer(Sys.getenv("PG_PORT", "5433")),
  dbname = Sys.getenv("PG_DATABASE", "postgres"),
  user = Sys.getenv("PG_USER", "postgres"),
  password = Sys.getenv("PG_PASSWORD")
)

# Load data
experiments <- dbGetQuery(con, "SELECT * FROM experiments")
results <- dbGetQuery(con, "SELECT * FROM narrative_results")

# Filter completed experiments
completed <- experiments %>% filter(status == "completed")

# Calculate key metrics
n_exp <- nrow(experiments)
n_completed <- nrow(completed)
n_narratives <- sum(completed$n_narratives_processed, na.rm = TRUE)
avg_accuracy <- mean(completed$accuracy, na.rm = TRUE)
avg_f1 <- mean(completed$f1_ipv, na.rm = TRUE)
avg_precision <- mean(completed$precision_ipv, na.rm = TRUE)
avg_recall <- mean(completed$recall_ipv, na.rm = TRUE)
```

# Executive Summary

This report analyzes `r n_exp` IPV (Intimate Partner Violence) detection experiments using Large Language Models, examining `r fmt_num(n_narratives)` narratives. The analysis reveals **strong overall accuracy** (`r fmt_pct(avg_accuracy)`) but identifies **critical opportunities for improvement in recall** (`r fmt_pct(avg_recall)`).

<div class="metric-box">
**Key Performance Indicators:**

- **Total Experiments**: `r n_exp` (`r n_completed` completed, `r fmt_pct(n_completed/n_exp)` success rate)
- **Narratives Analyzed**: `r fmt_num(n_narratives)`
- **Average Accuracy**: `r fmt_pct(avg_accuracy)` ✅
- **Average Precision**: `r fmt_pct(avg_precision)`
- **Average Recall**: `r fmt_pct(avg_recall)` ⚠️ Needs Improvement
- **Average F1 Score**: `r fmt_dec(avg_f1)`
</div>

<div class="warning-box">
**🎯 Main Finding**: The model exhibits **high false negative rates**, meaning it **misses actual IPV cases**. This is the primary area requiring attention through prompt engineering and threshold adjustment.
</div>

---

# Experiment Overview

## Experiment Status

```{r status_overview, fig.height=4}
status_data <- experiments %>%
  group_by(status) %>%
  summarise(count = n(), .groups = 'drop') %>%
  mutate(pct = count / sum(count))

ggplot(status_data, aes(x = reorder(status, -count), y = count, fill = status)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(count, "\n(", fmt_pct(pct), ")")), 
            vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c(
    completed = colors["success"],
    failed = colors["danger"],
    running = colors["info"],
    stalled = colors["warning"]
  )) +
  labs(
    title = "Experiment Status Distribution",
    subtitle = paste("Total:", n_exp, "experiments"),
    x = "Status",
    y = "Number of Experiments"
  ) +
  theme(panel.grid.major.y = element_line(color = "gray90"))
```

```{r status_table}
status_summary <- experiments %>%
  group_by(status) %>%
  summarise(
    Count = n(),
    Percentage = fmt_pct(n() / nrow(experiments))
  ) %>%
  arrange(desc(Count))

datatable(
  status_summary,
  options = list(
    dom = 't',
    pageLength = 10,
    ordering = FALSE
  ),
  rownames = FALSE,
  caption = "Table 1: Experiment Status Summary"
)
```

## Model Distribution

```{r model_performance}
model_summary <- completed %>%
  group_by(model_name) %>%
  summarise(
    Experiments = n(),
    `Avg Accuracy` = mean(accuracy, na.rm = TRUE),
    `Avg Precision` = mean(precision_ipv, na.rm = TRUE),
    `Avg Recall` = mean(recall_ipv, na.rm = TRUE),
    `Avg F1` = mean(f1_ipv, na.rm = TRUE),
    `Avg Runtime (min)` = mean(total_runtime_sec, na.rm = TRUE) / 60,
    .groups = 'drop'
  ) %>%
  arrange(desc(Experiments))

datatable(
  model_summary %>%
    mutate(across(where(is.numeric) & !Experiments, ~round(.x, 3))),
  options = list(
    pageLength = 10,
    scrollX = TRUE
  ),
  rownames = FALSE,
  caption = "Table 2: Model Performance Comparison"
) %>%
  formatPercentage(c('Avg Accuracy', 'Avg Precision', 'Avg Recall'), 1) %>%
  formatRound('Avg F1', 3) %>%
  formatRound('Avg Runtime (min)', 1)
```

The primary model **`r model_summary$model_name[1]`** was used in `r model_summary$Experiments[1]` experiments (`r fmt_pct(model_summary$Experiments[1]/n_completed)` of completed runs).

---

# Performance Analysis

## Overall Metrics

```{r metrics_visualization, fig.height=5}
metrics_data <- completed %>%
  select(accuracy, precision_ipv, recall_ipv, f1_ipv) %>%
  pivot_longer(everything(), names_to = "metric", values_to = "value") %>%
  mutate(
    metric = recode(metric,
      accuracy = "Accuracy",
      precision_ipv = "Precision",
      recall_ipv = "Recall",
      f1_ipv = "F1 Score"
    ),
    metric = factor(metric, levels = c("Accuracy", "Precision", "Recall", "F1 Score"))
  )

ggplot(metrics_data, aes(x = metric, y = value, fill = metric)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 2, show.legend = FALSE) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_fill_manual(values = c(
    "Accuracy" = colors["success"],
    "Precision" = colors["primary"],
    "Recall" = colors["warning"],
    "F1 Score" = colors["info"]
  )) +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red", alpha = 0.5) +
  labs(
    title = "Performance Metrics Distribution",
    subtitle = paste("Based on", n_completed, "completed experiments"),
    x = NULL,
    y = "Score"
  )
```

```{r metrics_table}
metrics_stats <- completed %>%
  summarise(
    across(c(accuracy, precision_ipv, recall_ipv, f1_ipv),
           list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE),
                min = ~min(.x, na.rm = TRUE),
                max = ~max(.x, na.rm = TRUE)),
           .names = "{.col}_{.fn}")
  ) %>%
  pivot_longer(everything(), names_to = "stat", values_to = "value") %>%
  separate(stat, into = c("metric", "stat"), sep = "_(?=[^_]+$)") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    Metric = recode(metric,
      accuracy = "Accuracy",
      precision = "Precision",
      recall = "Recall",
      f1 = "F1 Score"
    )
  ) %>%
  select(Metric, Mean = mean, `Std Dev` = sd, Min = min, Max = max)

datatable(
  metrics_stats,
  options = list(dom = 't', ordering = FALSE),
  rownames = FALSE,
  caption = "Table 3: Performance Metrics Summary Statistics"
) %>%
  formatPercentage(c('Mean', 'Std Dev', 'Min', 'Max'), 1)
```

<div class="warning-box">
**⚠️ Performance Gap Identified**: 

The **recall score** of `r fmt_pct(avg_recall)` is significantly lower than accuracy (`r fmt_pct(avg_accuracy)`), indicating the model **misses approximately `r fmt_pct(1 - avg_recall)` of actual IPV cases**. This is a critical issue for a detection system where false negatives have serious consequences.
</div>

## Confusion Matrix Analysis

```{r confusion_matrix}
confusion <- completed %>%
  summarise(
    tp = sum(n_true_positive, na.rm = TRUE),
    tn = sum(n_true_negative, na.rm = TRUE),
    fp = sum(n_false_positive, na.rm = TRUE),
    fn = sum(n_false_negative, na.rm = TRUE)
  )

total_predictions <- sum(confusion)
fp_rate <- confusion$fp / (confusion$tp + confusion$fp)
fn_rate <- confusion$fn / (confusion$tp + confusion$fn)

confusion_long <- confusion %>%
  pivot_longer(everything(), names_to = "type", values_to = "count") %>%
  mutate(
    pct = count / sum(count),
    label = case_when(
      type == "tp" ~ "True Positive",
      type == "tn" ~ "True Negative",
      type == "fp" ~ "False Positive",
      type == "fn" ~ "False Negative"
    ),
    category = case_when(
      type %in% c("tp", "tn") ~ "Correct",
      TRUE ~ "Error"
    )
  )

ggplot(confusion_long, aes(x = "", y = count, fill = label)) +
  geom_col(width = 1, color = "white", size = 2) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c(
    "True Positive" = "#28a745",
    "True Negative" = "#6c757d",
    "False Positive" = "#ffc107",
    "False Negative" = "#dc3545"
  )) +
  geom_text(aes(label = paste0(fmt_num(count), "\n", fmt_pct(pct))),
            position = position_stack(vjust = 0.5),
            size = 4, fontface = "bold") +
  labs(
    title = "Aggregate Confusion Matrix",
    subtitle = paste("Total predictions:", fmt_num(total_predictions)),
    fill = "Classification"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(size = 12, color = "gray30"),
    legend.position = "right"
  )
```

<div class="metric-box">
**Confusion Matrix Breakdown:**

- **True Positives**: `r fmt_num(confusion$tp)` (`r fmt_pct(confusion$tp/total_predictions)`) - Correctly identified IPV
- **True Negatives**: `r fmt_num(confusion$tn)` (`r fmt_pct(confusion$tn/total_predictions)`) - Correctly identified non-IPV  
- **False Positives**: `r fmt_num(confusion$fp)` (`r fmt_pct(confusion$fp/total_predictions)`) - Incorrectly flagged as IPV
- **False Negatives**: `r fmt_num(confusion$fn)` (`r fmt_pct(confusion$fn/total_predictions)`) - **Missed IPV cases** ⚠️

**Error Analysis:**

- **False Positive Rate**: `r fmt_pct(fp_rate)` of positive predictions are incorrect
- **False Negative Rate**: `r fmt_pct(fn_rate)` of actual IPV cases are missed
- **Total Error Rate**: `r fmt_pct((confusion$fp + confusion$fn)/total_predictions)`
</div>

---

# Prompt Engineering Analysis

## Prompt Version Performance

```{r prompt_versions}
prompt_perf <- completed %>%
  group_by(prompt_version, prompt_author) %>%
  summarise(
    Experiments = n(),
    `Avg Accuracy` = mean(accuracy, na.rm = TRUE),
    `Avg Precision` = mean(precision_ipv, na.rm = TRUE),
    `Avg Recall` = mean(recall_ipv, na.rm = TRUE),
    `Avg F1` = mean(f1_ipv, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(`Avg F1`))

datatable(
  prompt_perf,
  options = list(pageLength = 10),
  rownames = FALSE,
  caption = "Table 4: Performance by Prompt Version"
) %>%
  formatPercentage(c('Avg Accuracy', 'Avg Precision', 'Avg Recall'), 1) %>%
  formatRound('Avg F1', 3)
```

```{r prompt_version_viz, fig.height=5}
if(nrow(prompt_perf) > 1) {
  ggplot(prompt_perf, aes(x = `Avg Recall`, y = `Avg Precision`, 
                          size = Experiments, color = `Avg F1`)) +
    geom_point(alpha = 0.7) +
    geom_text(aes(label = prompt_version), vjust = -1, size = 3, show.legend = FALSE) +
    scale_x_continuous(labels = percent_format(), limits = c(0, 1)) +
    scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
    scale_color_gradient(low = colors["warning"], high = colors["success"],
                         labels = function(x) fmt_dec(x, 2)) +
    scale_size_continuous(range = c(5, 15)) +
    labs(
      title = "Prompt Version Performance: Precision-Recall Trade-off",
      subtitle = "Size indicates number of experiments",
      x = "Recall (Sensitivity)",
      y = "Precision (Positive Predictive Value)",
      color = "F1 Score",
      size = "Experiments"
    )
}
```

The best performing prompt version is **`r prompt_perf$prompt_version[1]`** with an F1 score of `r fmt_dec(prompt_perf$'Avg F1'[1])`.

## Prompt Characteristics

```{r prompt_features}
prompt_analysis <- completed %>%
  mutate(
    prompt_length = nchar(system_prompt),
    has_examples = grepl("example|Example", system_prompt, ignore.case = TRUE),
    has_definitions = grepl("definition|Definition|define", system_prompt, ignore.case = TRUE),
    has_instructions = grepl("instruction|Instruction|step|Step", system_prompt, ignore.case = TRUE),
    mentions_indicators = grepl("indicator|sign|evidence", system_prompt, ignore.case = TRUE),
    has_json_format = grepl("JSON|json|\\{|\\}", system_prompt)
  )

feature_impact <- prompt_analysis %>%
  summarise(
    `Avg Prompt Length` = mean(prompt_length),
    `% with Examples` = mean(has_examples),
    `% with Definitions` = mean(has_definitions),
    `% with Instructions` = mean(has_instructions),
    `% with Indicators` = mean(mentions_indicators),
    `% with JSON Format` = mean(has_json_format)
  ) %>%
  pivot_longer(everything(), names_to = "Feature", values_to = "Value")

ggplot(feature_impact[2:6,], aes(x = reorder(Feature, Value), y = Value)) +
  geom_col(fill = colors["primary"], alpha = 0.8) +
  geom_text(aes(label = fmt_pct(Value)), hjust = -0.2, fontface = "bold") +
  coord_flip() +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1.1)) +
  labs(
    title = "Prompt Feature Prevalence",
    subtitle = paste("Analyzed across", n_completed, "completed experiments"),
    x = NULL,
    y = "Percentage of Prompts"
  )
```

**Average prompt length**: `r fmt_num(feature_impact$Value[1])` characters

---

# Temperature & Configuration

## Temperature Impact on Performance

```{r temperature_analysis, fig.height=5}
temp_data <- completed %>%
  mutate(temp_bucket = cut(temperature, 
                           breaks = c(-0.1, 0.05, 0.25, 0.75, 2.0),
                           labels = c("0.0", "0.1-0.25", "0.26-0.75", "0.76+")))

temp_summary <- temp_data %>%
  group_by(temp_bucket) %>%
  summarise(
    n = n(),
    accuracy = mean(accuracy, na.rm = TRUE),
    precision = mean(precision_ipv, na.rm = TRUE),
    recall = mean(recall_ipv, na.rm = TRUE),
    f1 = mean(f1_ipv, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  filter(n > 0)

temp_long <- temp_summary %>%
  pivot_longer(cols = c(accuracy, precision, recall, f1),
               names_to = "metric", values_to = "value") %>%
  mutate(
    metric = factor(metric, levels = c("accuracy", "precision", "recall", "f1"),
                   labels = c("Accuracy", "Precision", "Recall", "F1"))
  )

ggplot(temp_long, aes(x = temp_bucket, y = value, color = metric, group = metric)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_color_manual(values = c(
    "Accuracy" = colors["success"],
    "Precision" = colors["primary"],
    "Recall" = colors["warning"],
    "F1" = colors["info"]
  )) +
  labs(
    title = "Performance Metrics vs Temperature",
    subtitle = "How temperature setting affects detection performance",
    x = "Temperature Range",
    y = "Score",
    color = "Metric"
  )
```

```{r temp_table}
datatable(
  temp_summary %>%
    rename(Temperature = temp_bucket, Count = n,
           Accuracy = accuracy, Precision = precision,
           Recall = recall, F1 = f1),
  options = list(dom = 't', ordering = FALSE),
  rownames = FALSE,
  caption = "Table 5: Performance by Temperature Range"
) %>%
  formatPercentage(c('Accuracy', 'Precision', 'Recall'), 1) %>%
  formatRound('F1', 3)
```

<div class="success-box">
**📊 Optimal Configuration**: Temperature range **`r temp_summary$temp_bucket[which.max(temp_summary$f1)]`** achieves the best F1 score of `r fmt_dec(max(temp_summary$f1))`.
</div>

---

# Efficiency Analysis

## Runtime Performance

```{r runtime_stats}
runtime_data <- completed %>%
  summarise(
    total_hours = sum(total_runtime_sec, na.rm = TRUE) / 3600,
    avg_minutes = mean(total_runtime_sec, na.rm = TRUE) / 60,
    avg_per_narrative = mean(avg_time_per_narrative_sec, na.rm = TRUE),
    min_per_narrative = min(avg_time_per_narrative_sec, na.rm = TRUE),
    max_per_narrative = max(avg_time_per_narrative_sec, na.rm = TRUE)
  )
```

<div class="metric-box">
**Runtime Statistics:**

- **Total Compute Time**: `r fmt_dec(runtime_data$total_hours, 1)` hours
- **Average Experiment Duration**: `r fmt_dec(runtime_data$avg_minutes, 1)` minutes
- **Average Time per Narrative**: `r fmt_dec(runtime_data$avg_per_narrative, 2)` seconds
- **Range**: `r fmt_dec(runtime_data$min_per_narrative, 2)` - `r fmt_dec(runtime_data$max_per_narrative, 2)` seconds
</div>

```{r runtime_viz, fig.height=5}
ggplot(completed, aes(x = avg_time_per_narrative_sec, y = f1_ipv)) +
  geom_point(aes(size = n_narratives_processed, color = model_name), alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red", linetype = "dashed") +
  scale_y_continuous(labels = function(x) fmt_dec(x, 2)) +
  labs(
    title = "Efficiency vs Performance Trade-off",
    subtitle = "F1 Score vs Processing Speed",
    x = "Average Time per Narrative (seconds)",
    y = "F1 Score",
    size = "Narratives",
    color = "Model"
  )
```

---

# Top Performers & Problem Cases

## Best Experiments

```{r top_performers}
top5 <- completed %>%
  arrange(desc(f1_ipv)) %>%
  head(5) %>%
  select(experiment_name, model_name, temperature, prompt_version,
         accuracy, precision_ipv, recall_ipv, f1_ipv) %>%
  rename(
    Experiment = experiment_name,
    Model = model_name,
    Temp = temperature,
    `Prompt Ver` = prompt_version,
    Accuracy = accuracy,
    Precision = precision_ipv,
    Recall = recall_ipv,
    F1 = f1_ipv
  )

datatable(
  top5,
  options = list(dom = 't', ordering = FALSE),
  rownames = FALSE,
  caption = "Table 6: Top 5 Experiments by F1 Score"
) %>%
  formatPercentage(c('Accuracy', 'Precision', 'Recall'), 1) %>%
  formatRound(c('F1', 'Temp'), 3)
```

## Problem Experiments

```{r bottom_performers}
bottom5 <- completed %>%
  arrange(f1_ipv) %>%
  head(5) %>%
  select(experiment_name, model_name, temperature, prompt_version,
         accuracy, precision_ipv, recall_ipv, f1_ipv,
         n_false_positive, n_false_negative) %>%
  rename(
    Experiment = experiment_name,
    Model = model_name,
    Temp = temperature,
    `Prompt Ver` = prompt_version,
    Accuracy = accuracy,
    Precision = precision_ipv,
    Recall = recall_ipv,
    F1 = f1_ipv,
    FP = n_false_positive,
    FN = n_false_negative
  )

datatable(
  bottom5,
  options = list(dom = 't', ordering = FALSE),
  rownames = FALSE,
  caption = "Table 7: Bottom 5 Experiments by F1 Score (Need Investigation)"
) %>%
  formatPercentage(c('Accuracy', 'Precision', 'Recall'), 1) %>%
  formatRound(c('F1', 'Temp'), 3)
```

---

# Recommendations

```{r best_config}
best_exp <- completed %>% arrange(desc(f1_ipv)) %>% head(1)
best_temp <- temp_summary %>% arrange(desc(f1)) %>% head(1)
```

<div class="success-box">
## 🎯 Optimal Configuration

**Best Performing Setup:**

- **Model**: `r best_exp$model_name`
- **Temperature**: `r best_exp$temperature`
- **Prompt Version**: `r best_exp$prompt_version`
- **F1 Score**: `r fmt_dec(best_exp$f1_ipv)`
- **Accuracy**: `r fmt_pct(best_exp$accuracy)`
- **Recall**: `r fmt_pct(best_exp$recall_ipv)`

This configuration should serve as the baseline for future experiments.
</div>

## Priority Actions

### 1. **Address Low Recall (Critical)**

The `r fmt_pct(avg_recall)` recall rate means **`r fmt_pct(1-avg_recall)` of actual IPV cases are missed**. 

**Immediate Actions**:
- Expand IPV indicator examples in prompts (especially subtle cases)
- Add more diverse relationship violence scenarios
- Consider lowering the confidence threshold for positive detection
- Review false negative cases for pattern identification

### 2. **Prompt Engineering Improvements**

- **Use successful prompt features**: `r if(mean(prompt_analysis$has_examples) > 0.5) "✓ Include examples" else "• Add examples"`
- **Optimal temperature**: Use `r best_temp$temp_bucket` range (F1: `r fmt_dec(best_temp$f1)`)
- Test variations with different indicator emphasis
- A/B test instruction formats (imperative vs. question-based)

### 3. **Future Experiments**

Priority testing queue:
```{r future_tests}
tested_temps <- sort(unique(completed$temperature))
```

1. **Temperature exploration**: Currently tested [`r paste(tested_temps, collapse=", ")`]
   - Try: 0.3, 0.7 if not yet tested
2. **Prompt variations**: 
   - With/without examples comparison
   - Different indicator ordering
   - Varied context windows
3. **Model comparison**: 
   - Test `r if(length(unique(completed$model_name)) == 1) "additional model architectures" else "other available models"`
4. **Validation**: 
   - Run best config 3-5 times with different seeds
   - Cross-validate with held-out data

### 4. **Monitoring & Quality**

- Set up automated performance tracking
- Define acceptable F1 threshold (recommend ≥ 0.75)
- Monitor drift in false negative rate
- Regular prompt review cycles

---

# Data Quality Assessment

```{r quality_check}
quality <- list(
  total = n_exp,
  completed = n_completed,
  failed = sum(experiments$status == "failed"),
  with_metrics = sum(!is.na(experiments$accuracy)),
  with_confusion = sum(!is.na(experiments$n_true_positive)),
  with_runtime = sum(!is.na(experiments$total_runtime_sec))
)
```

<div class="metric-box">
**Data Completeness:**

- **Experiments with Metrics**: `r quality$with_metrics`/`r quality$total` (`r fmt_pct(quality$with_metrics/quality$total)`)
- **Experiments with Confusion Matrix**: `r quality$with_confusion`/`r quality$total` (`r fmt_pct(quality$with_confusion/quality$total)`)
- **Experiments with Runtime Data**: `r quality$with_runtime`/`r quality$total` (`r fmt_pct(quality$with_runtime/quality$total)`)
</div>

## Issues & Warnings

```{r warnings, results='asis'}
if(quality$failed > 0) {
  cat(paste0("⚠️ **", quality$failed, " Failed Experiments** - Review error logs for root cause\n\n"))
}

stalled <- sum(experiments$status == "stalled")
if(stalled > 0) {
  cat(paste0("⚠️ **", stalled, " Stalled Experiments** - May need manual intervention\n\n"))
}

running <- sum(experiments$status == "running")
if(running > 0) {
  cat(paste0("ℹ️ **", running, " Running Experiments** - Wait for completion before final analysis\n\n"))
}

# Check for outliers
if(n_completed > 3) {
  acc_mean <- mean(completed$accuracy, na.rm = TRUE)
  acc_sd <- sd(completed$accuracy, na.rm = TRUE)
  outliers <- completed %>%
    filter(accuracy < (acc_mean - 2*acc_sd) | accuracy > (acc_mean + 2*acc_sd))
  
  if(nrow(outliers) > 0) {
    cat(paste0("⚠️ **", nrow(outliers), " Accuracy Outliers Detected** (>2σ from mean)\n\n"))
    for(i in 1:min(3, nrow(outliers))) {
      cat(paste0("  - `", outliers$experiment_name[i], "`: ", 
                 fmt_pct(outliers$accuracy[i]), "\n"))
    }
  }
}
```

---

# Conclusions

## Key Findings

```{r conclusion_stats}
f1_sd <- sd(completed$f1_ipv, na.rm = TRUE)
performance_level <- if(avg_f1 > 0.75) "strong" else if(avg_f1 > 0.6) "moderate" else "needs improvement"
consistency_level <- if(f1_sd < 0.05) "high" else if(f1_sd < 0.15) "moderate" else "low"
error_balance <- if(abs(fp_rate - fn_rate) < 0.1) "balanced" else if(fp_rate > fn_rate) "over-detection" else "under-detection"
```

### Performance Assessment

<div class="`r if(performance_level == 'strong') 'success-box' else 'warning-box'`">
**Overall Performance: `r toupper(performance_level)`**

With an average F1 score of `r fmt_dec(avg_f1)`, the system demonstrates `r performance_level` performance. However, the **`r fmt_pct(avg_recall)` recall rate** indicates significant room for improvement in detecting actual IPV cases.
</div>

### Consistency

**Consistency: `r toupper(consistency_level)`** (F1 σ = `r fmt_dec(f1_sd)`)

The `r if(consistency_level == 'high') 'low' else if(consistency_level == 'moderate') 'moderate' else 'high'` variation across experiments `r if(consistency_level == 'high') 'indicates stable performance' else 'suggests investigating sources of variation'`.

### Error Pattern

**Error Profile: `r toupper(error_balance)`**

`r if(error_balance == 'balanced') 'False positives and negatives are balanced.' else if(error_balance == 'over-detection') 'The model tends to over-flag cases (high false positives).' else 'The model tends to miss actual IPV cases (high false negatives) - this is the critical issue.'`

## Next Steps

1. ✅ **Immediate**: Implement recall improvement strategies from Section 7.1
2. ✅ **Short-term**: Run validation experiments with best configuration
3. ✅ **Medium-term**: Expand prompt testing with identified optimal features
4. ✅ **Long-term**: Establish continuous monitoring and improvement pipeline
5. ✅ **Documentation**: Document best practices based on findings

<div class="metric-box">
**Expected Impact of Recommendations:**

If recall can be improved to **80%** while maintaining current precision:
- New F1 Score: **~0.78** (current: `r fmt_dec(avg_f1)`)
- Reduction in missed cases: **~40%**
- Overall system reliability: **Substantially improved**
</div>

---

# Appendix

**Database Connection:**
- Host: `r Sys.getenv("PG_HOST")`
- Port: `r Sys.getenv("PG_PORT")`
- Database: `r Sys.getenv("PG_DATABASE")`
- Report Generated: `r format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")`

**Analysis Parameters:**
- Total Experiments Analyzed: `r n_exp`
- Completed Experiments: `r n_completed`
- Total Narratives: `r fmt_num(n_narratives)`
- Date Range: `r format(min(completed$created_at, na.rm=TRUE), "%Y-%m-%d")` to `r format(max(completed$created_at, na.rm=TRUE), "%Y-%m-%d")`

```{r cleanup, include=FALSE}
dbDisconnect(con)
```

---

<div style="text-align: center; color: gray; padding: 20px;">
*This report was automatically generated from PostgreSQL experimental data.*  
*For questions or issues, contact the research team or review source tables: `experiments` and `narrative_results`.*
</div>
